---
title: "R Intro"
---

Load R packages.

```{r}
#| message: false
library(tidyverse)
library(sf)
library(tmap)
```

## Read CSV

Read GBIF CSV.

```{r}
#| message: false
beaver_df <- read_csv('data/raw/Castor_canadensis-12106.csv')
```

Preview GBIF data.

```{r}
glimpse(beaver_df)
```

Update GBIF data so that it can plotted on a map. `st_as_sf()` will create a vector point for each record using decimalLongitude and decimalLatitude, add a geometry column, and set the coordinate reference system.

```{r}
beaver_sf <- st_as_sf(beaver_df, 
                      coords = c('decimalLongitude', 'decimalLatitude'), 
                      crs=4326, 
                      remove=FALSE)
```

## Read GeoJSON

Read a GeoJSON file that has the boundaries for Oregon state.

```{r}
oregon_sf <- read_sf('data/cleaned/oregon.geojson')
```

Check if GBIF data and Oregon boundary have same CRS.

```{r}
st_crs(oregon_sf) == st_crs(beaver_sf)
```
Update CRS for Oregon boundaries.

```{r}
oregon_sf <- st_transform(oregon_sf, crs=st_crs(beaver_sf))
st_crs(oregon_sf) == st_crs(beaver_sf)
```


## Interactive maps

Create interactive map.

```{r}
tmap_mode("view")
```

Create map with GBIF data and Oregon boundary. Color code each GBIF occurrence record using the basis field. Use Open Street Maps as basemap.

```{r}
tm_shape(oregon_sf) +
  tm_lines() +
  tm_shape(beaver_sf) +
  tm_symbols(fill='basis') +
  tm_basemap("OpenStreetMap")
```

## Static maps

Create static map.

```{r}
tmap_mode("plot")
```
Create map with GBIF data and Oregon boundary. Color code each GBIF occurrence record using the basis field. Use Open Street Maps as basemap.

```{r}
tm_shape(oregon_sf) +
  tm_lines() +
  tm_shape(beaver_sf) +
  tm_symbols(fill='basis') +
  tm_basemap("OpenStreetMap")
  
```

Update map appearance. Add a north arrow, scale bar, and a title.

```{r}
map <- tm_shape(oregon_sf) +
  tm_lines() +
  tm_shape(beaver_sf) +
  tm_symbols(fill='basis') +
  tm_basemap("OpenStreetMap")  +
  tm_compass(position = c("left", "bottom")) +
  tm_scalebar(position = c("left", "bottom")) +
  tm_title("Beaver occurrence in Oregon")  +
  tm_layout(inner.margins = c(0.1,.1),
            legend.position=c("right", "bottom"),
            legend.bg.color = 'white')

map
```

Save map as png image.

```{r}
#| eval: false
tmap_save(map, filename = "output/beaver_oregon.png", width = 7, height = 7)

```

## Bar charts

Create bar chart. Use basis values for the x axis.

```{r}
ggplot(beaver_df) +
  geom_bar(mapping=aes(x=basis))
```

Update chart appearance. Add color to the bars, a title, and custom theme.

```{r}
chart <- ggplot(beaver_df) +
  geom_bar(mapping=aes(x=basis), fill='75cd5e')  + 
  labs(title = 'Beaver occurrence in Oregon')  +
  theme_bw() +
  theme(title = element_text(size = 14), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
  
chart
```

Save chart as png image.

```{r}
#| eval: false
ggsave('output/chart.png', chart, width = 5, height=5)
```
